---
title: "Prediction of Increase in Reaction Time with use of cell phones while driving"
author: "Vikas Vig"
date: "November 25, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Overview

This was built as part of assigment submission for the course Developing Data Products as part of the Coursera Data Science Specialization.  

The submission is in two parts:  
- a Shiny Application (https://vigvikas.shinyapps.io/ReactionTime/)  
- Reproducible Pitch Presentation (this document)  

The Shiny Appliction is built based on **"reaction.time"** dataset from R package "Using R". This dataset uses simulated data based on values from the NHTSA tudy "The Influence of the Use of Mobile Phones on Driver Situation Awareness"   

**A simple application is considered and built to demonstrate Shiny App development.**

## Exploring Dataset
Structure of Dataset "reaction.time"
```{r echo = FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(UsingR)
library(dplyr)
library(gridExtra)
data("reaction.time")

myDB <- reaction.time
str(myDB)
```
Data frame has 60 observations on the following 4 variables:  
1. age with two factors: 16-24 or 25+;   
2. gender with two factors: M or F;   
3. control with two factors: T = using cell phone or C; and   
4. time in seconds  

## Graphically looking at changes in Reaction Time 

```{r echo = FALSE, warning=FALSE, message=FALSE}
p <- ggplot(myDB, aes(x = gender, y = time, color = control))
p <- p + geom_boxplot()

q <- ggplot(myDB, aes(x = age, y = time, color = control))
q <- q + geom_boxplot()

grid.arrange(p, q, ncol=1)
```

## Model Building
Since the reaction time (both with and without use of cell phones) seem to be influenced by Age & Gender a simple Regression Model was chosen. Two regression models as below were developed.   
**No use of cell phones**
```{r echo = FALSE, warning=FALSE, message=FALSE}
dbControl <- filter(myDB, control == "C")
dbTel <- filter(myDB, control == "T")
regModCon <- lm(time ~ age + gender, data = dbControl)
regModCon$coefficients
```
**Use of cell phones**
```{r echo = FALSE, warning=FALSE, message=FALSE}
regModTel <- lm(time ~ age + gender, data = dbTel)
regModTel$coefficients
```
**These models were used to develop the Shiny App.The user is required to enter Age & Gender and the app displays increase in reaction time due to use of cell phone while driving.**
